---
title: "Project Report"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
library(readxl)
library(patchwork)
library(leaflet)
library(rvest)
library(plotly)
library(gganimate)
library(png)
knitr::opts_chunk$set(
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r, include=FALSE}
songs_abuse=read_csv("data/songs_abuse.csv")
genly_billboard=read_csv("data/genly_billboard.csv")
```

<style type="text/css">
</style>
&nbsp;

## Motivation

This project is overall motivated by the current major public health issue in United States: the Illicit drug use. Previous studies showed a relationship between illicit drug use and pop music [[1](https://pubmed.ncbi.nlm.nih.gov/30599351/),[2](https://www.tandfonline.com/doi/full/10.3109/10826084.2012.637433)], indicating that pop music may have a potential encouraging effect on illicit drug use [[3](https://pubmed.ncbi.nlm.nih.gov/18250243/)]. However, these studies are either cross-sectional, or need to be reproduce using more recent data. 
In this study, we will describe the trend of music genre, frequency of illicit drug in the popular music, and the relationship between popular music and illicit drug use during the period of 1999-2019.

&nbsp;

## Questions

The questions for this project are based on the 2 datasets we used which are explained in the next section:

<li>What are the trends of substance use among our interested age group at national level and state level during the period of 1999-2019?</li>
<li>What are the trends of popular music genre and description of substance use in the lyrics during the period of 1999-2019?</li>
<li>Is there any association exist between popular music and substance use during the period of 1999-2019?</li>

&nbsp;
## Data 

